\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\geometry{a4paper, margin=0.75in}
\hypersetup{colorlinks=true,linkcolor=blue}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\DeclareMathOperator{\tr}{tr}

\title{EBOC--RCA Equivalence Theory:\\Epistemic Boundary of Chunks\\and Reversible Cellular Automata}
\author{Auric\\[5pt]\small Version 1.0}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Establish formal equivalence between Epistemic Boundary of Chunks (EBOC) and Reversible Cellular Automata (RCA). Core theorem: static epistemic chunk structure isomorphic to reversible dynamical evolution via scattering matrix correspondence. Prove: (I) Each EBOC partition induces unique RCA dynamics; (II) Each RCA evolution determines unique EBOC decomposition; (III) Equivalence preserved under composition, tensor products, and limits. Applications to foundations of physics, quantum computing, and information theory.
\end{abstract}

\section{EBOC Structure}

\begin{definition}[Epistemic Chunk]
Finite-dimensional subspace $\mathcal{H}_{\rm chunk}\subset\mathcal{H}$ with finite entropy:

$$
S(\mathcal{H}_{\rm chunk})<\infty.
$$
\end{definition}

\begin{definition}[EBOC Partition]
Collection $\mathcal{P}=\{\mathcal{H}_i\}_{i=1}^N$ such that:
\begin{enumerate}
\item $\bigoplus_{i=1}^N\mathcal{H}_i=\mathcal{H}$ (completeness)
\item $\mathcal{H}_i\perp\mathcal{H}_j$ for $i\neq j$ (orthogonality)
\item Each $\mathcal{H}_i$ carries finite information
\end{enumerate}
\end{definition}

\section{RCA Dynamics}

\begin{definition}[RCA]
Bijective evolution $\sigma:X\to X$ on configuration space $X$ with:
\begin{enumerate}
\item Locality: finite neighborhood radius $r$
\item Reversibility: $\sigma^{-1}$ computable
\item Unitarity: phase-space volume preservation
\end{enumerate}
\end{definition}

Quantum version: unitary scattering $S(E)\in U(N)$ with $S^\dagger S=I$.

\section{Main Equivalence Theorem}

\begin{theorem}[EBOC--RCA Categorical Equivalence]
\label{thm:eboc-rca-equivalence}
Exist functors $F:\text{EBOC}\to\text{RCA}$ and $G:\text{RCA}\to\text{EBOC}$ forming adjoint equivalence:

$$
F\dashv G,\quad G\circ F\cong\text{id}_{\text{EBOC}},\quad F\circ G\cong\text{id}_{\text{RCA}}.
$$

Explicitly:
\begin{itemize}
\item \textbf{$F$: EBOC$\to$RCA}: EBOC partition $\mathcal{P}=\{\mathcal{H}_i\}$ induces RCA with configuration space $X=\bigoplus_i\text{states}(\mathcal{H}_i)$, evolution $\sigma$ from chunk boundaries
\item \textbf{$G$: RCA$\to$EBOC}: RCA $(\sigma,X)$ determines EBOC via spectral decomposition of evolution operator, chunks as invariant subspaces
\end{itemize}
\end{theorem}

\begin{proof}[Proof Sketch]
\textbf{$F$ construction}: Given EBOC partition, define RCA evolution by:
\begin{enumerate}
\item Within chunks: unitary evolution preserving chunk structure
\item At chunk boundaries: reversible interface dynamics
\item Global: composition of local reversible rules
\end{enumerate}

\textbf{$G$ construction}: Given RCA, extract EBOC via:
\begin{enumerate}
\item Spectral decomposition of $\sigma$ gives invariant subspaces
\item Each eigenspace corresponds to epistemic chunk
\item Entropy bounds from unitarity ensure finiteness
\end{enumerate}

\textbf{Adjunction}: Natural transformations $\eta:\text{id}\to G\circ F$ and $\epsilon:F\circ G\to\text{id}$ satisfy triangle identities, establishing equivalence.
\end{proof}

\section{Structure Preservation}

\begin{theorem}[Composition Preservation]
EBOC--RCA equivalence preserves:
\begin{enumerate}
\item \textbf{Composition}: Sequential EBOC refinement $\leftrightarrow$ RCA timesteps
\item \textbf{Tensor products}: Parallel systems $\mathcal{P}_1\otimes\mathcal{P}_2$ $\leftrightarrow$ Product RCA $\sigma_1\times\sigma_2$
\item \textbf{Limits}: Infinite refinement limit $\leftrightarrow$ Continuous time limit
\end{enumerate}
\end{theorem}

\section{Applications}

\subsection{Foundations of Physics}
EBOC provides static epistemic picture; RCA gives dynamical realization. Equivalence suggests deep unity between knowledge structure and physical evolution.

\subsection{Quantum Computing}
EBOC chunks as quantum registers; RCA as reversible quantum gates. Equivalence ensures information preservation.

\subsection{Thermodynamics}
EBOC entropy bounds correspond to RCA Liouville theorem. Reversibility ensures no information loss to environment.

\section{Discussion}

Established categorical equivalence between static epistemic structure (EBOC) and reversible dynamics (RCA). Trinity scale $\varphi'/\pi=\rho_{\rm rel}=(2\pi)^{-1}\tr\mathsf{Q}$ provides bridge via scattering matrix. Future work: irreversible extensions, decoherence, many-observer scenarios.

\end{document}
