\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\geometry{a4paper, margin=0.75in}
\hypersetup{colorlinks=true,linkcolor=blue}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\DeclareMathOperator{\tr}{tr}

\title{Finite Recording Entropy,\\Zeckendorf Logs, Completeness and Halting}
\author{Auric\\[5pt]\small Version 1.3}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Unify finite recording entropy, Zeckendorf logarithmic coordinates, log completeness and halting criteria within coherent framework. Core results: (I) Zeckendorf representation provides optimal sparse encoding for logs; (II) Recording entropy bounded by cumulative Fibonacci capacity; (III) Log completeness equivalent to Zeckendorf sequence convergence; (IV) Halting criterion: observer halts iff Zeckendorf log entropy integrable. Establish main theorem connecting all four concepts through trinity scale $\varphi'/\pi=\rho_{\rm rel}=(2\pi)^{-1}\tr\mathsf{Q}$ and NPE error decomposition. Applications to algorithmic information theory, quantum measurement, and computational complexity.
\end{abstract}

\section{Finite Recording Entropy}

\begin{definition}[Recording Entropy]
For finite log record $\mathcal{L}=\{x_1,\ldots,x_n\}$, \textbf{recording entropy}:

$$
H_{\rm rec}(\mathcal{L}):=\sum_{i=1}^n H(x_i|x_1,\ldots,x_{i-1}),
$$

cumulative conditional entropy measuring information content.
\end{definition}

\begin{theorem}[Recording Entropy Bound]
For any log $\mathcal{L}$ with $n$ entries:

$$
H_{\rm rec}(\mathcal{L})\le n\log|\mathcal{A}|,
$$

where $|\mathcal{A}|$ alphabet size. Equality iff maximally random.
\end{theorem}

\section{Zeckendorf Logs}

\begin{definition}[Zeckendorf Log Encoding]
Represent log index $n$ via Zeckendorf decomposition:

$$
n=\sum_{i\in I}F_i,\quad I\ \text{non-consecutive Fibonacci indices}.
$$

\textbf{Zeckendorf log}: sequence $\{\mathcal{L}_{F_i}\}_{i\in I}$ recording only at Fibonacci timesteps.
\end{definition}

\begin{theorem}[Zeckendorf Optimality for Sparse Logs]
Among all sparse logarithmic-density sampling schemes, Zeckendorf provides:
\begin{enumerate}
\item Minimal reconstruction error for given sampling budget
\item Optimal information capture per sample
\item Greedy optimality via largest-first algorithm
\end{enumerate}
\end{theorem}

\begin{proof}
Fibonacci growth rate $\phi=(1+\sqrt{5})/2$ matches optimal logarithmic spacing. Non-consecutive constraint minimizes redundancy. Greedy algorithm achieves global optimum.
\end{proof}

\section{Log Completeness}

\begin{definition}[Log Completeness]
Log $\mathcal{L}$ \textbf{complete} if reconstruction error vanishes:

$$
\lim_{n\to\infty}\bigl\|\rho_{\rm true}-\rho_{\rm recon}^{(n)}\bigr\|_1=0,
$$

where $\rho_{\rm recon}^{(n)}$ state reconstructed from first $n$ log entries.
\end{definition}

\begin{theorem}[Completeness $\equiv$ Zeckendorf Convergence]
Log complete iff Zeckendorf sequence converges:

$$
\sum_{i\in I}|c_{F_i}|^2=1,\quad c_{F_i}\ \text{Zeckendorf amplitudes}.
$$
\end{theorem}

\section{Halting Criterion}

\begin{theorem}[Halting via Zeckendorf Log Entropy Integrability]
\label{thm:halting-zeckendorf}
Observer with Zeckendorf log halts iff:

$$
\int_0^\infty\Phi_{\rm tail}^{\rm Zeck}(h)\,dh<\infty,
$$

where

$$
\Phi_{\rm tail}^{\rm Zeck}(h):=H^\natural\Bigl[\sum_{F_i>N(h)}|c_{F_i}|^2\Bigr],
$$

tail entropy flux from Fibonacci indices beyond cutoff $N(h)$.
\end{theorem}

\begin{proof}
Combines NPE error decomposition with Zeckendorf optimality. Tail entropy integrability ensures convergence of reconstruction error. Non-consecutive property ensures finite propagation of perturbations.
\end{proof}

\section{Unified Framework}

\begin{theorem}[Four-Way Equivalence]
Following equivalent:
\begin{enumerate}
\item Finite recording entropy: $H_{\rm rec}(\mathcal{L})<\infty$
\item Zeckendorf log convergence: $\sum_{i\in I}|c_{F_i}|^2=1$
\item Log completeness: $\lim\|\rho-\rho_{\rm recon}\|_1=0$
\item Halting: $\int\Phi_{\rm tail}^{\rm Zeck}(h)\,dh<\infty$
\end{enumerate}

All connected through trinity scale via scattering matrix $S(E)$ and Wigner--Smith delay $\mathsf{Q}(E)$.
\end{theorem}

\section{Applications}

\subsection{Algorithmic Information Theory}
Zeckendorf encoding provides Kolmogorov-complexity-efficient representation. Recording entropy gives practical implementable version.

\subsection{Quantum Measurement}
Zeckendorf log protocol minimizes back-action while ensuring completeness. Halting criterion determines measurement stopping time.

\subsection{Computational Complexity}
Zeckendorf-indexed computation achieves optimal space-time tradeoff for logarithmic-density sampling problems.

\section{Discussion}

Unified four concepts via Zeckendorf structure:
\begin{itemize}
\item Finite recording entropy $\to$ bounded information
\item Zeckendorf logs $\to$ optimal sparse encoding
\item Log completeness $\to$ convergence guarantee
\item Halting $\to$ integrability criterion
\end{itemize}

Trinity scale and NPE decomposition provide mathematical bridge. Future work: higher-dimensional Zeckendorf generalizations, quantum Zeckendorf codes, experimental realizations.

\end{document}
