\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\geometry{a4paper, margin=0.75in}
\hypersetup{colorlinks=true,linkcolor=blue}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\DeclareMathOperator{\tr}{tr}

\title{WLRC Theory:\\Windowed Logs, Reversible Computation\\and Resource-Constrained Halting}
\author{Auric\\[5pt]\small Version 1.0}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Establish theory connecting Windowed Logs, Reversible Computation (RC) and Resource-Constrained halting. Core framework: (I) Windowed log system records finite-bandwidth observations with NPE error decomposition; (II) Reversible computation preserves information via bijective dynamics; (III) Resource constraints (time, space, energy) impose fundamental bounds on log completeness; (IV) Halting criterion: observer halts iff windowed log entropy flux integrable under resource bounds. Prove main theorem establishing equivalence between log completeness, reversible computation halting, and resource-constrained decidability.
\end{abstract}

\section{Windowed Log System}

\begin{definition}[Windowed Log]
\textbf{Windowed log} $\mathcal{L}_w$ finite-bandwidth recording system:

$$
\mathcal{L}_w(t):=\int w(E-E_0)\,\rho(E,t)\,dE,
$$

where $w$ window function, $\rho(E,t)$ time-dependent spectral density.
\end{definition}

NPE error decomposition:

$$
\varepsilon_{\rm log}(h)=\varepsilon_{\rm alias}(h)+\varepsilon_{\rm EM}(h)+\varepsilon_{\rm tail}(h).
$$

\section{Reversible Computation}

Bijective dynamics $\sigma:X\to X$ with $\sigma^{-1}$ computable. Scattering matrix $S(E)$ unitary: $S^\dagger S=I$.

Information preservation via Liouville theorem (phase-space volume) or unitary evolution (Hilbert space norm).

\section{Resource Constraints}

\begin{definition}[Resource Triple]
$(T,M,E)$ where $T$ time, $M$ space (memory), $E$ energy.
\end{definition}

\begin{theorem}[Resource-Constrained Halting]
Observer with resource bound $(T_{\max},M_{\max},E_{\max})$ halts iff:

\begin{enumerate}
\item Time: $\int_0^{T_{\max}}\Phi_{\rm tail}(t)\,dt<\infty$
\item Space: $\dim\mathcal{H}_{\rm log}\le M_{\max}$
\item Energy: $\int E\,d\mu_{\rm log}(E)\le E_{\max}$
\end{enumerate}

where $\Phi_{\rm tail}$ tail entropy flux, $\mathcal{H}_{\rm log}$ log Hilbert space, $\mu_{\rm log}$ log spectral measure.
\end{theorem}

\section{Main Theorem}

\begin{theorem}[WLRC Equivalence]
Following equivalent:
\begin{enumerate}
\item Windowed log complete within tolerance
\item Reversible computation halts under resource bounds
\item Tail entropy flux integrable
\item NPE error vanishes in refined scale limit
\end{enumerate}
\end{theorem}

\section{Discussion}

Unified windowed logs, reversible computation and resource-constrained halting through entropy flux integrability criterion. Applications to bounded-error quantum computation, thermodynamic computing, and finite-resource decidability theory.

\end{document}
